Symmetric Circuits for Rank Gates

### 2018-03-26 sections/symmetric-circuits.tex
Minor revisions to Sections 3 and 4.

As our aim is a circuit characterisation of $\FPR$, whose expressive power is
strictly greater than that of $\FPC$, we need to consider gates corresponding to
non-symmetric Boolean functions. In particular, we consider \emph{rank} gates
whose inputs are a matrix and whose result is invariant under row and column
permutations. To lead up to this, we first develop a general framework of
structured Boolean functions. These are functions whose inputs naturally encode
$\tau$-structures and where the output is invariant under the natural symmetries
of such structures. The\emph{matrix-invariant} functions are a natural
partiruclar case. We therefore define symmetric circuits in a general form where
gates can be labelled by \emph{isomorphism-invariant} structured functions.

Ultimately, our aim is replicate the results Anderson and
Dawar~\cite{AndersonD17} in this more general setting. A difficulty we face is
that the proof methods in that paper heavily rely on the assumption that the
Boolean functions computed by individual gates are symmetric functions. We
address this difficulty in the next section.

@@

% AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau := (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s
  \in S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the trivial relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol $s$ in
$\tau$ to the set $D_s$ and each relation symbol in $\tau$ to the trivial
relation over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and
denote it by $\str{\tau, D}$.

@@

% GW: move sorted permutations into the background section
For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$. Let $X := \ind (\tau, D)$ and
$G \leq \sym_{X}$. We note that $G$ there is an obvious group action of $G$ on
$X$. Let $\aut (\tau, D)$ be the automorphism group of $\str{\tau, D}$. We
define a group action of $\aut (\tau, D)$ on $X$ for $\sigma \in \aut(\tau, D)$
by $\sigma \cdot \vec{a}_R := (\sigma \vec{a})_R$, for all $\vec{a}_R \in X$. We
identify $\aut(\tau, D)$ with the subgroup of $\sym_X$ that acts equivalently on
$X$.

@@

Structured functions give us a general framework for indexing the input sequence
to a Boolean function and associating with that function an appropriate symmetry
condition, or notion of equivalence. Let $F$ be a $(\tau, D)$-structured
function. We say $F$ is \emph{$G$-invariant} if it is constant on $G$-equivalent
input vectors, i.e.\ for all $G$-equivalent $f, g \in \{0,1\}^{X}$, $F(f) =
F(g)$. We say $F$ is \emph{isomorphism-invariant} if it is $\aut(\tau,
D)$-invariant. Note that a structured function with vocabulary $\tau$ is
isomorphism-invariant if, and only if, it decides a property of
$\tau$-structures.

@@

\begin{definition}[Circuits on Structures]
  Let $\mathbb{B}$ be a basis and $\rho$ be a relational vocabulary, we define a
  \emph{$(\mathbb{B}, \rho)$-circuit} $C$ of order $n$ computing a $q$-ary query
  $Q$ as a structure $\langle G, \Omega, \Sigma, \Lambda, L \rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C$ and $\vert C \vert := \vert G
    \vert$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B} \uplus \rho \uplus
    \{0,1\} $ such that $\vert \Sigma^{-1} (0) \vert \leq 1$ and $\vert
    \Sigma^{-1} (1) \vert \leq 1$. Those gates mapped to $\rho \uplus \{0,1\}$
    are called input gates, with those mapped to $\rho$ called relational gates
    and those mapped $\{0,1\}$ called constant gates. Those gates mapped to
    $\mathbb{B}$ are called internal gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_{R_i})_{R_i \in
      R}$ such that $\Lambda_{R_i}$ maps each relational gate $g$ with $\Sigma
    (g) = R_i$ to the tuple $\Lambda_{R_i} (g) \in [n]^{r}$. When no ambiguity
    arises we write $\Lambda (g)$ for $\Lambda_{R_i} (g)$.
  \item $L$ associates with each internal gate $g$ a function $L(g):
    \ind(\Sigma(g)) \rightarrow G$ such that if we define a relation $W
    \subseteq G^{2}$ by $W(h_1,h_2)$ iff $h_2$ is an internal gate and $h_1$ is
    in the image of $L(h_2)$, then $(G, W)$ is a directed acyclic graph.
  \end{itemize}
\end{definition}

@@

Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
\rho)$-circuit of order $n$. The order of a circuit denotes the fact that it
takes as input structures of size $n$. If $(C_n)_{n \in \nats}$ is a family of
circuits we assume that each $C_n$ is a circuit of order $n$.

@@

  We now present a number of observations, and then combine these observations
  to prove both directions of the hypothesis. Let $\mathcal{B}$ be a
  $\rho$-structure of size $n$ over the universe $[n]$. We can define a function
  $f_{\mathcal{B}} : \ind(\rho, [n]) \rightarrow {0,1}$ such that
  $f_{\mathcal{B}} (\vec{a}_R) = 1$ if, and only if, $\vec{a} \in
  R^{\mathcal{B}}$. Let $f : \ind(\rho, [n]) \rightarrow \{0,1\}$. We can define
  the $\rho$-structure $\mathcal{B}_f$ over the universe $[n]$ such that for all
  $R \in \rho$, we have $\vec{a} \in R^{\mathcal{B}_f}$ if, and only if,
  $f(\vec{a}_R) = 1$. It is easy to see that the functions $\mathcal{B} \mapsto
  f_{\mathcal{B}}$ and $f \mapsto \mathcal{B}_f$ are inverse to one another, and
  so define a bijection. Moreover, we notice that for all $\sigma \in \sym_n$ we
  have that $f_{\sigma \mathcal{B}}(\vec{a}_R) = 1$ if, and only if, $\vec{a}
  \in R^{\sigma \mathcal{B}}$ if, and only if, $\sigma^{-1} \vec{a} \in
  R^{\mathcal{B}}$ if, and only if, $(f_{\mathcal{B}} \sigma^{-1})(\vec{a}_R) =
  f_{\mathcal{B}}(\sigma^{-1} \vec{a}_R)= 1$. It follows $f_{\sigma \mathcal{B}}
  = f_{\mathcal{B}} \sigma^{-1}$.

  Let $\mathcal{A}$ be a $\rho$-structure of size $n$ over the universe $U$, and
  let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$. From the definition of $F_C$
  it follows that for $\gamma \in \sym_n$ we have that $F_C (f_{\gamma
    \mathcal{A}})$ if, and only if, $C[\gamma \mathcal{A}](g_{\text{o}}) = 1$.

@@

\begin{definition}[Automorphism]
  let $C = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B},\tau)$-circuit of order $n$ computing a $q$-ary, and where
  $\mathbb{B}$ is a basis of isomorphism-invariant structured functions. Let
  $\sigma \in \sym_n$ and $\pi: G \rightarrow G$ be a bijection such that
  \begin{myitemize}
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item For each pair of gates $g, h \in G$ $W(h,g)$ if and only if $W(\pi h,
    \pi g)$ and for each internal gate $g$ we have that $L(\pi g)$ and $ \pi
    \cdot L(g)$ are isomorphism-equivalent.
  \end{myitemize}
  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{extends to an automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut (C)$.
\end{definition}

@@

It follows that for any symmetric circuit $C$ of order $n$ there is a
homomorphism $h$ that maps $\sym_n$ to $\aut(C)$ such that if $\sigma \in
\sym_n$ then $h(\sigma)$ is an an automorphism extending $\sigma$. Suppose $C$
does not contain a relational gate labelled by a relation symbol with non-zero
arity. In that case $C$ computes a constant function. As such, in this paper we
always assume a circuit contains at least one relational gate with non-zero
arity. As such by assumption there exists a relational gate in $C$ such that
some element of $[n]$ appears in the tuple labelling that gate. By symmetry it
follows that every elements of $[n]$ appears in a tuple labelling a relational
gate in $C$. It follows that no two distinct elements of $\sym_n$ agree on all
input gates, and so the homomorphism $h$ is injective.

@@

In order to make use of these technical tools, as well ensure the
polynomial-time decidability of many important circuit properties, we should
like to be able to construct a normal form analogous to rigidity and present a
polynomial time algorithm that transforms a circuit into an equivalent circuit
of this form. It is at this point that we arrive at the first complication
introduced by our generalisation. Anderson and Dawar's polynomial-time
translation makes indispensable of the polynomial-time decidability of many
important circuit properties for circuits defined over symmetric bases. However,
for the more general circuits discussed here, it is not known if even the most
basic circuit properties are polynomial-time decidable. Indeed, we show in
Section~\ref{sec:transparent} that for circuits that include gates computing
non-symmetric functions all of the relevant circuit properties of interest are
at least as hard to decide as the graph-isomorphism problem. As such,
constructing an analogous argument, as well as establishing the numerous other
crucial results whose proofs rely on the polynomial-time decidability of various
circuit properties, would be beyond the scope of this paper.

@@

Before we can formally define \emph{transparency} we need to define the
\emph{syntactic-equivalence} relation on the gates of a circuit. The intuition
is that two gates in a circuit are syntactically-equivalent if the circuits
underneath them are `hereditary equivalent', i.e.\ the two circuits underneath
these gates are really just copies of one another.
% In order to prove this result, Anderson and Dawar introduce the notion of a
% \emph{rigid} circuit, proving that symmetric circuits may be

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. We recursively define the equivalence relation
  \emph{syntactic-equivalence}, which we denote using the symbol `$\equiv$', on
  $G$ as follows. Suppose $g$ and $h$ are gates in $C$ such that $\Sigma (g) =
  \Sigma(h)$ and either both $g$ and $h$ are output gates and $\Omega^{-1}(g) =
  \Omega^{-1}(h)$ or neither are output gates. Suppose $g$ and $h$ are input
  gates, then $g \equiv h$ if, and only if, both $g$ and $h$ are constant gates
  or both are relational gates and $\Lambda(g) = \Lambda (h)$. Suppose $g$ and
  $h$ are internal gates and and suppose we have defined the
  syntactic-equivalence relation for all gates of depth less than the depth of
  either $g$ or $h$. Then $g \equiv h$ if, and only if, $L(g) /_\equiv$ and
  $L(h) /_\equiv$ are isomorphism-equivalent.
\end{definition}

@@

For a circuit $C$ of order $n$ and a gate $g$ we map think of $g$ as computing
the function that maps an input structure $\mathcal{A}$ and a bijection $\gamma$
from the universe of $\mathcal{A}$ to $[n]$ to the evaluation $C[\gamma
\mathcal{A}(g)$. While we would like to be able to identify gates that compute
the same function in this sense, it is not hard to show that deciding this
equivalence relation for a given circuit is $\NP$-hard. We now show that if two
gates are syntactically-equivalent then the functions computed at these two
gates must be equal. We show later that the syntactic-equivalence relation is
polynomial-time decidable for the class of circuits of interest to us in this
paper. In this sense we shall treat syntactic-equivalence as a tractable
refinement of the $\NP$-complete relation.

@@

\begin{lem}
  Let $C$ be a circuit of order $n$, $\sigma \in \sym_n$ and $\pi, \pi' \in
  \aut(C)$ both extend $\sigma$, then for every gate $g$ in the circuit we have
  that $\pi (g)$ and $\pi'(g)$ are syntactically-equivalent.
  \label{lem:permutation-extending-syntactic-equivalence}
\end{lem}
\begin{proof}
  We first note that, from the definition of an automorphism, for any gate $g$
  in $C$, $\Sigma (g) = \Sigma (\pi (g)) = \Sigma (\pi' (g))$, and either all of
  $g$, $\pi (g)$ and $\pi (g')$ are output gates and $\pi \Omega
  (\Omega^{-1}(g)) = \Omega (\sigma \Omega^{-1}(g)) = \pi' \Omega
  (\Omega^{-1})$, or none of $g$, $\pi(g)$ or $\pi' (g)$ are output gates.

@@

\begin{definition}
  Let $C$ be a circuit and $g$ be a gate in $C$. We say $g$ has \emph{injective
    labels} if $L(g)$ is an injection. We say $g$ has \emph{unique children} if
  no two gates in $H_g$ are syntactically-equivalent. We say $g$ has
  \emph{unique labels} if $g$ has injective labels and unique children. We say
  $C$ has \emph{injective labels} (resp. \emph{unique labels}) if every gate in
  $C$ has injective labels (resp. unique labels). We say $C$ is
  \emph{transparent} if every gate $g$ in $C$ such that $\Sigma(g)$ is not a
  symmetric function has unique labels.
\end{definition}

It is worth noting that if a circuit has injective labels and is reduced then
then it has unique labels. The converse is, in general, false.

There is a small technical point that needs some discussion before we continue.
The definition of a circuit allows for the inclusion of a gate such that there
is no path from this gate to any of the output gates. In other words, we can be
sure, not matter the rest of the circuit, that any such gate cannot have any
effect the function computed by the circuit. We say these gates are
\emph{redundant}.

@@

It is easy to see that we can construct from a given circuit an equivalent
circuit without redundant gates by simply removing all of the redundant gates.
Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
$n$ and let $g$ be a redundant circuit in $C$. It is easy to see that $g$ is not
the child of a non-redundant gate, nor is it a relational or output gate. Thus
we can define the circuit $C' := \langle G', \Omega, \restr{\Sigma}{G'},
\Lambda, \restr{L}{G'} \rangle$, where $G' \subseteq G$ is the set of all
non-redundant gates in $C$.

@@

We have from Proposition~\ref{prop:unique-children-unique-extensions} that
circuits with unique gates have unique extensions, but, in order to prove many
of the results we will need later, we not only need that each automorphism is
uniquely defined by the permutation it extends, but also that the isomorphisms
that witness the isomorphism-equivalences, and so witness the validity of the
automorphism, are also uniquely defined (or, at least very least, well
constrained) by the permutation $\sigma$. Without a sufficient constraint on the
choice of isomorphism, testing if two gates have isomorphism-equivalent
labellings (a requirement for computing the automorphism extending a
permutation) requires solving some version of the isomorphism problem for
$\tau$-structures, where $\tau$ is the vocabulary of the gates in question.
Indeed, we prove in Section~\ref{sec:transparent}, that for the class of
circuits with unique children deciding if there is an isomorphism-equivalence
between two gates is as hard as the graph-isomorphism problem. It follows that a
great many important circuit properties (e.g. the orbit of a gate, symmetry of a
circuit, etc.) for circuits with unique gates are at the very least not
obviously polynomial-time decidable. As such, circuits with unique children,
although satisfying the requirement of having unique extensions, are
insufficient for our purposes.

@@

\begin{prop}
  \label{prop:fuctions-maj}
  There is a deterministic algorithm that outputs for each symmetric function
  $F: \{0,1\}^n \rightarrow \{0,1\}$ (encoded as a binary $n$-tuple as above) a
  symmetric circuit $C$ defined over the basis $\BM$ that computes $F$.
  Moreover, this function is computable in time polynomial in $n$ and the
  circuit $C$ has depth at most $5$, width at most $2n+2$ and size at most $5n +
  3$.
\end{prop}

>>>>

As our ultimate aim is a circuit characterisation of $\FPR$, whose expressive power is
strictly greater than that of $\FPC$, we need to consider gates corresponding to
non-symmetric Boolean functions. In particular, we consider \emph{rank} gates
whose inputs are a matrix and whose result is invariant under row and column
permutations. To lead up to this, we first develop a general framework of
structured Boolean functions. These are functions whose inputs naturally encode
$\tau$-structures and where the output is invariant under the natural symmetries
of such structures. The \emph{matrix-invariant} functions (defined
formally below) are a natural
particular case. We therefore define symmetric circuits in a general form where
gates can be labelled by \emph{isomorphism-invariant} structured functions.

Ultimately, our aim is to replicate the results Anderson and
Dawar~\cite{AndersonD17} in this more general setting. A difficulty we face is
that the proof methods in that paper heavily rely on the assumption that the
Boolean functions computed by individual gates are symmetric functions. We
address this difficulty in the next section.

@@

% AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau := (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s
  \in S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the full relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol $s$ in
$\tau$ to the set $D_s$ and each relation symbol in $\tau$ to the full
relation over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and
denote it by $\str{\tau, D}$.

@@

% GW: move sorted permutations into the background section
For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$. Let $X := \ind (\tau, D)$ and
$G \leq \sym_{X}$. We note that there is an obvious group action of $G$ on
$X$. Let $\aut (\tau, D)$ be the automorphism group of $\str{\tau, D}$. We
define a group action of $\aut (\tau, D)$ on $X$ for $\sigma \in \aut(\tau, D)$
by $\sigma \cdot \vec{a}_R := (\sigma \vec{a})_R$, for all $\vec{a}_R \in X$. We
identify $\aut(\tau, D)$ with the subgroup of $\sym_X$ that acts equivalently on
$X$.

@@

Structured functions give us a general framework for indexing the input sequence
to a Boolean function and associating with that function an appropriate symmetry
condition, or notion of equivalence. Let $F$ be a $(\tau, D)$-structured
function. We say $F$ is \emph{$G$-invariant} if it is constant on $G$-equivalent
input vectors, i.e.\ for all $G$-equivalent $f, g \in \{0,1\}^{X}$, $F(f) =
F(g)$. We say $F$ is \emph{isomorphism-invariant} if it is $\aut(\tau,
D)$-invariant. Note that a structured function with vocabulary $\tau$ is
isomorphism-invariant if, and only if, it determines a property of
$\tau$-structures.

@@

\begin{definition}[Circuits on Structures]
  Let $\mathbb{B}$ be a basis and $\rho$ be a relational vocabulary, we define a
  \emph{$(\mathbb{B}, \rho)$-circuit} $C$ of order $n$ computing a $q$-ary query
  $Q$ as a structure $\langle G, \Omega, \Sigma, \Lambda, L \rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C$.% and $\vert C \vert := \vert G    \vert$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B} \uplus \rho \uplus
    \{0,1\} $ such that $\vert \Sigma^{-1} (0) \vert \leq 1$ and $\vert
    \Sigma^{-1} (1) \vert \leq 1$. Those gates mapped to $\rho \uplus \{0,1\}$
    are called input gates, with those mapped to $\rho$ called relational gates
    and those mapped to $\{0,1\}$ called constant gates. Those gates mapped to
    $\mathbb{B}$ are called internal gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_{R_i})_{R_i \in
      R}$ such that $\Lambda_{R_i}$ maps each relational gate $g$ with $\Sigma
    (g) = R_i$ to the tuple $\Lambda_{R_i} (g) \in [n]^{r}$. When no ambiguity
    arises we write $\Lambda (g)$ for $\Lambda_{R_i} (g)$.
  \item $L$ associates with each internal gate $g$ a function $L(g):
    \ind(\Sigma(g)) \rightarrow G$ such that if we define a relation $W
    \subseteq G^{2}$ by $W(h_1,h_2)$ iff $h_2$ is an internal gate and $h_1$ is
    in the image of $L(h_2)$, then $(G, W)$ is a directed acyclic graph.
  \end{itemize}
\end{definition}

The definition requires some explanation.  Each gate in $G$ computes a
function of its inputs and the relation $W$ on $G$ is the set of
``wires''.  That is, $W(h,g)$ indicates that the value computed at $h$
is an input to $g$.  However, since the functions are structured, we
need more information on the set of inputs to $g$ and this is provided
by the labelling $L$.  $\Sigma(g)$ tells us what the function computed
at $g$ is, and thus $\ind(\Sigma(g))$ tells us the structure on the
inputs and $L(g)$ maps this to the set of gates that form the inputs
to $g$.

@@

Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
\rho)$-circuit of order $n$.  Recall that the orde of a circuit refers
to the size of the input structures it takes.  In contrast, we define
the \emph{size} of $C$, denoted $\vert C \vert$ to be the number of
elements in $G$. If $(C_n)_{n \in \nats}$ is a family of
circuits we assume that each $C_n$ is a circuit of order $n$.

@@

  We now present a number of observations, and then combine these observations
  to prove both directions of the lemma. Let $\mathcal{B}$ be a
  $\rho$-structure of size $n$ over the universe $[n]$. We can define a function
  $f_{\mathcal{B}} : \ind(\rho, [n]) \rightarrow {0,1}$ such that
  $f_{\mathcal{B}} (\vec{a}_R) = 1$ if, and only if, $\vec{a} \in
  R^{\mathcal{B}}$. Let $f : \ind(\rho, [n]) \rightarrow \{0,1\}$. We can define
  the $\rho$-structure $\mathcal{B}_f$ over the universe $[n]$ such that for all
  $R \in \rho$, we have $\vec{a} \in R^{\mathcal{B}_f}$ if, and only if,
  $f(\vec{a}_R) = 1$. It is easy to see that the functions $\mathcal{B} \mapsto
  f_{\mathcal{B}}$ and $f \mapsto \mathcal{B}_f$ are inverse to one another, and
  so define a bijection. Moreover, we notice that for all $\sigma \in \sym_n$ we
  have that $f_{\sigma \mathcal{B}}(\vec{a}_R) = 1$ if, and only if, $\vec{a}
  \in R^{\sigma \mathcal{B}}$ if, and only if, $\sigma^{-1} \vec{a} \in
  R^{\mathcal{B}}$ if, and only if, $(f_{\mathcal{B}} \sigma^{-1})(\vec{a}_R) =
  f_{\mathcal{B}}(\sigma^{-1} \vec{a}_R)= 1$. It follows $f_{\sigma \mathcal{B}}
  = f_{\mathcal{B}} \sigma^{-1}$.

  Let $\mathcal{A}$ be a $\rho$-structure of size $n$ over the universe $U$, and
  let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$. From the definition of $F_C$
  it follows that for $\gamma \in \sym_n$ we have that $F_C (f_{\gamma
    \mathcal{A}})=1$ if, and only if, $C[\gamma \mathcal{A}](g_{\text{o}}) = 1$.

@@

\begin{definition}[Automorphism]\label{defn:automorphism}
  Let $C = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B},\tau)$-circuit of order $n$ computing a $q$-ary query, and where
  $\mathbb{B}$ is a basis of isomorphism-invariant structured functions. Let
  $\sigma \in \sym_n$ and $\pi: G \rightarrow G$ be a bijection such that
  \begin{myitemize}
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item For each pair of gates $g, h \in G$ $W(h,g)$ if and only if $W(\pi h,
    \pi g)$ and for each internal gate $g$ we have that $L(\pi g)$ and $ \pi
    \cdot L(g)$ are isomorphism-equivalent.
  \end{myitemize}
  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{extends to an automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut (C)$.
\end{definition}

@@

It follows that for any symmetric circuit $C$ of order $n$ there is a
homomorphism $h$ that maps $\sym_n$ to $\aut(C)$ such that if $\sigma \in
\sym_n$ then $h(\sigma)$ is an an automorphism extending $\sigma$.  Suppose $C$
does not contain a relational gate labelled by a relation symbol with non-zero
arity.  In that case $C$ computes a constant function.  For this reason, in this paper we
always assume a circuit contains at least one relational gate with non-zero
arity.  Now, by assumption there exists a relational gate in $C$ such that
some element of $[n]$ appears in the tuple labelling that gate.  By symmetry it
follows that every elements of $[n]$ appears in a tuple labelling a relational
gate in $C$. It follows that no two distinct elements of $\sym_n$ agree on all
input gates, and so the homomorphism $h$ is injective.

@@

In order to make use of these technical tools, as well ensure the
polynomial-time decidability of many important circuit properties, we should
like to be able to construct a normal form analogous to rigidity and present a
polynomial time algorithm that transforms a circuit into an equivalent circuit
of this form.  It is at this point that we arrive at the first complication
introduced by our generalisation. The polynomial-time
translation in~\cite{AndersonD17} makes indispensable of the polynomial-time decidability of many
important circuit properties for circuits defined over symmetric
bases.  However,
for the more general circuits discussed here, it is not known if even the most
basic circuit properties are polynomial-time decidable.  This is
essentially due to the requirement built in to
Definition~\ref{defn:automorphism} that an automorphism that takes $g$
to $g'$ must be an isomorphism between $L(g)$ and $L(g')$, which makes
checking the condition as hard as isomorphism checking.
Indeed, we show in
Section~\ref{sec:transparent} that for circuits that include gates computing
non-symmetric functions all of the relevant circuit properties of interest are
at least as hard to decide as the graph-isomorphism problem. As such,
constructing an argument analogous to~\cite{AndersonD17}, as well as establishing the numerous other
crucial results whose proofs rely on the polynomial-time decidability of various
circuit properties, would be beyond the scope of this paper.

@@

Before we can formally define \emph{transparency} we need to define the
\emph{syntactic-equivalence} relation on the gates of a circuit. The intuition
is that two gates in a circuit are syntactically-equivalent if the circuits
underneath them are `hereditarily equivalent', i.e.\ the two circuits underneath
these gates are really just copies of one another.
% In order to prove this result, Anderson and Dawar introduce the notion of a
% \emph{rigid} circuit, proving that symmetric circuits may be

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. We recursively define the equivalence relation
  \emph{syntactic-equivalence}, which we denote using the symbol `$\equiv$', on
  $G$ as follows. Suppose $g$ and $h$ are gates in $C$ such that $\Sigma (g) =
  \Sigma(h)$ and either both $g$ and $h$ are output gates and $\Omega^{-1}(g) =
  \Omega^{-1}(h)$ or neither are output gates. Suppose $g$ and $h$ are input
  gates, then $g \equiv h$ if, and only if, both $g$ and $h$ are constant gates
  or both are relational gates and $\Lambda(g) = \Lambda (h)$.  Suppose $g$ and
  $h$ are internal gates and suppose we have defined the
  syntactic-equivalence relation for all gates of depth less than the depth of
  either $g$ or $h$. Then $g \equiv h$ if, and only if, $L(g) /_\equiv$ and
  $L(h) /_\equiv$ are isomorphism-equivalent.
\end{definition}

@@

For a circuit $C$ of order $n$ and a gate $g$ we think of $g$ as computing
the function that maps an input structure $\mathcal{A}$ and a bijection $\gamma$
from the universe of $\mathcal{A}$ to $[n]$ to the evaluation $C[\gamma
\mathcal{A}(g)]$. While we would like to be able to identify gates that compute
the same function in this sense, it is not hard to show that deciding this
equivalence relation for a given circuit is $\NP$-hard. We now show that if two
gates are syntactically-equivalent then the functions computed at these two
gates must be equal. We show later that the syntactic-equivalence relation is
polynomial-time decidable for the class of circuits of interest to us in this
paper. In this sense we shall treat syntactic-equivalence as a tractable
refinement of the $\NP$-complete relation.

@@

\begin{lem}
  Let $C$ be a circuit of order $n$, $\sigma \in \sym_n$ and $\pi, \pi' \in
  \aut(C)$ both extend $\sigma$, then for every gate $g$ in the circuit we have
  that $\pi (g)$ and $\pi'(g)$ are syntactically-equivalent.
  \label{lem:permutation-extending-syntactic-equivalence}
\end{lem}
\begin{proof}
  We first note that, from the definition of an automorphism, for any gate $g$
  in $C$, $\Sigma (g) = \Sigma (\pi (g)) = \Sigma (\pi' (g))$, and either all of
  $g$, $\pi (g)$ and $\pi (g')$ are output gates and $\pi (g) = \pi \Omega
  (\Omega^{-1}(g)) = \Omega (\sigma \Omega^{-1}(g)) = \pi' \Omega
  (\Omega^{-1}(g)) = \pi'(g)$, or none of $g$, $\pi(g)$ or $\pi' (g)$ are output gates.

@@

\begin{definition}
  Let $C$ be a circuit and $g$ be a gate in $C$. We say $g$ has \emph{injective
    labels} if $L(g)$ is an injection. We say $g$ has \emph{unique children} if
  no two gates in $H_g$ are syntactically-equivalent. We say $g$ has
  \emph{unique labels} if $g$ has injective labels and unique children. We say
  $C$ has \emph{injective labels} (resp.\ \emph{unique labels}) if every gate in
  $C$ has injective labels (resp.\ unique labels). We say $C$ is
  \emph{transparent} if every gate $g$ in $C$ such that $\Sigma(g)$ is not a
  symmetric function has unique labels.
\end{definition}

It is worth noting that if a circuit has injective labels and is reduced 
then it has unique labels. The converse is, in general, false.

There is a small technical point that needs some discussion before we continue.
The definition of a circuit allows for the inclusion of a gate such that there
is no path from this gate to any of the output gates. In other words, we can be
sure, no matter the rest of the circuit, that any such gate cannot have any
effect the function computed by the circuit. We say these gates are
\emph{redundant}.

@@

It is easy to see that we can construct from a given circuit an equivalent
circuit without redundant gates by simply removing all of the redundant gates.
Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
$n$ and let $g$ be a redundant gate in $C$.  It is easy to see that $g$ is not
the child of a non-redundant gate, nor is it a relational or output gate. Thus
we can define the circuit $C' := \langle G', \Omega, \restr{\Sigma}{G'},
\Lambda, \restr{L}{G'} \rangle$, where $G' \subseteq G$ is the set of all
non-redundant gates in $C$.

@@

We have from Proposition~\ref{prop:unique-children-unique-extensions} that
circuits with unique gates have unique extensions, but, in order to prove many
of the results we will need later, we not only need that each automorphism is
uniquely defined by the permutation it extends, but also that the isomorphisms
that witness the isomorphism-equivalences, and so witness the validity of the
automorphism, are also uniquely defined ( or, at the very least, well
constrained) by the permutation $\sigma$. Without a sufficient constraint on the
choice of isomorphism, testing if two gates have isomorphism-equivalent
labellings (a requirement for computing the automorphism extending a
permutation) requires solving some version of the isomorphism problem for
$\tau$-structures, where $\tau$ is the vocabulary of the gates in question.
Indeed, we prove in Section~\ref{sec:transparent}, that for the class of
circuits with unique children deciding if there is an isomorphism-equivalence
between two gates is as hard as the graph-isomorphism problem. It follows that a
great many important circuit properties (e.g.\ the orbit of a gate, symmetry of a
circuit, etc.) for circuits with unique gates are at the very least not
obviously polynomial-time decidable.  As such, circuits with unique children,
although satisfying the requirement of having unique extensions, are
insufficient for our purposes.

@@

\begin{prop}
  \label{prop:fuctions-maj}
  There is a deterministic algorithm that outputs for each symmetric function
  $F: \{0,1\}^n \rightarrow \{0,1\}$ (encoded as a binary $n$-tuple as above) a
  symmetric circuit $C$ defined over the basis $\BM$ that computes $F$.
  Moreover, this algorithm runs in time polynomial in $n$ and the
  circuit $C$ has depth at most $5$, width at most $2n+2$ and size at most $5n +
  3$.
\end{prop}
